Reasoning Graphically about Quantum Computation

### 2008-02-28 papers/reasoning-graphically/paper.tex
working...
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1387 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[runningheads]{llncs}
%\usepackage[dvips]{color}
%\usepackage{multicol}
%\usepackage{tabls}
%\usepackage{ttbox}
\usepackage{allmtt}
%\usepackage{pstricks,pst-node,pst-tree} % PS Tricks for diagrams
%\usepackage{proof2}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{stmaryrd}
%\usepackage{mbboard}
\usepackage{xspace}
\usepackage{diagrams}

@@

$$\minterp{G} = (\vinterp{G'}\;|\; G' \in \binterp{G})$$

@@

Rules that have been proved for concrete graphs, such as the
Frombenius Law, are in fact a family of proofs that correspond to
replacing the boundary edges with variable nodes. This is beacuse of
the symmetry of rules in a compact closed categories:

@@

In this section we show how the formalism we have presented for graph
patterns can be used to provide a \emph{meta-level} framework for
reasoning about a given model of a compact closed
categories. Following the terminology of logical frameworks such as
Isabelle~\cite{isabelle}, we call the specification provided by the
underlying model an \emph{object-level} graph formalism. Each
object-level formalism defines a set of rules which are treated as the
axioms for the system. The object-level implementation also defines
the data at the nodes and edges as well as corresponding data-matching
behaviour.

These are not true of the graphs abstractly, but are proved of the
particular model of concrete graphs being considered.

@@

A rule in our framework is a pair of pattern graph that represent the
left and right hand sides of an equation. Substituon with a rule then
involves replacing a subgraph that matches the left hand side with the
right hand side. To do this there has to be a isomorphism between
variable nodes in the left and right hand subgraphs. More concretely,
there must be a unique naming of the variable nodes in the left hand
side such that each name occurs in a unique variable node on the right
hand side. Given a matching on the left hand side, the target subgraph
can be replaced with right hand side connected by the corresponding
variable nodes. This provides rewriting for pattern graphs.

\subsubsection{Soundness.} An object-level graph system is sound if
their is some pair of graphs which can not be derived as an equation.

To ensure soundness, it is sufficient to show that the object-level
model is a compact-closed category and that the basic equational rules
of the object-level can be proved in the underlying model.

\subsubsection{Completeness.} A formalism for graphs can be shown to
be complete with respect to the underlying semantics if the equational
rules have been shown to be complete.

\subsection{Implications}

@@

The model we are interested in is finite dimensional hilbert spaces as
an account of qantum computation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[runningheads]{llncs}
%\usepackage[dvips]{color}
%\usepackage{multicol}
%\usepackage{tabls}
%\usepackage{ttbox}
\usepackage{allmtt}
%\usepackage{pstricks,pst-node,pst-tree} % PS Tricks for diagrams
%\usepackage{proof2}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{stmaryrd}
%\usepackage{mbboard}
\usepackage{xspace}
\usepackage{diagrams}
\usepackage{proof2}

@@

$$\minterp{G} = \{\vinterp{G'}\;.\; G' \in \binterp{G}\}$$

@@

@@

In this section we outline how the formalism we have presented for
graph patterns can be used to provide a \emph{meta-level} framework
for reasoning about a given model of a compact closed
categories. Following the terminology of logical frameworks such as
Isabelle~\cite{isabelle}, we call the specification provided by the
underlying model an \emph{object-level} graph formalism. Each
object-level formalism defines a set of rules which are treated as the
axioms for the system. The object-level implementation also defines
the data at the nodes and edges as well as corresponding data-matching
behaviour.

@@

A rule in our framework declarative in the sense that it is a pair of
pattern graph that represent the left and right hand sides of an
equation. Substituon with a rule then involves replacing a subgraph
that matches the left hand side with the right hand side. To do this
there has to be a isomorphism between variable nodes in the left and
right hand subgraphs. More concretely, there must be a unique naming
of the variable nodes in the left hand side such that each name occurs
in a unique variable node on the right hand side. Given a matching on
the left hand side, the target subgraph can be replaced with right
hand side connected by the corresponding variable nodes. This provides
rewriting for pattern graphs.

\subsection{Meta-Level Logic and Derived Rules}

The meta-logical rules of the framework are quite simple as they only
involve dealing with object-level equations:

\begin{center}
\prooftree
A = B \in \Gamma
\justifies
\Gamma \vdash A = B
\using\mbox{trivial}
\endprooftree
\quad\quad
\prooftree
\justifies
\Gamma \vdash A = A
\using\mbox{reflexivity}
\endprooftree
\end{center}

\begin{center}
\prooftree
\Gamma \vdash A = B
\justifies
\Gamma \vdash B = A
\using\mbox{symmetry}
\endprooftree
\quad\quad
\prooftree
\Gamma \vdash A = B
\Gamma \vdash C = D
\justifies
\Gamma \vdash (C = D[A/B])
\using\mbox{substitution}
\endprooftree
\end{center}

\noindent where $\Gamma$ is the set of object-level axioms. For the
reflexivity rule, $A$ must be a well-formed pattern graph.

This system allows derived rules which allow a combination of steps to
be abbreviated by representing them as a single new rule.

\subsection{Soundness and Completeness}

The soundness of a system that uses our meta-level rests on two
arguments:

\begin{itemize}
\item Representation of the underying model as a compact closed
  category with a defined set of equations.

\item The soundness and adequacy of the axioms when expressed within
  the compact closed category. Adquency ensures that the encoding as a
  graph pattern represents only the intended rules.  The soundness of
  the rules is obviously needed and requires proving the rule in terms
  of the underlying semantics.
\end{itemize}

The completeness of the resulting system, with respect to the intended
semantics, depends on the completeness of the set of given rules. The
completeness of

\subsection{Derived Rules}

\subsection{Simplification}

% An object-level graph system is sound if their is some pair of graphs
% which can not be derived as an equation.

% To ensure soundness, it is sufficient to show that the object-level
% model is a compact-closed category and that the basic equational rules
% of the object-level can be proved in the underlying model.

% The axioms should be proved of the particular model of concrete graphs
% being considered.

% \subsubsection{Completeness.} A formalism for graphs can be shown to
% be complete with respect to the underlying semantics if the equational
% rules have been shown to be complete.

% \subsection{Implications}

@@

The model of qantum computation in which we are interested is finite
dimensional hilbert spaces.
